---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# gt3x_comparison

<!-- badges: start -->
<!-- badges: end -->

The goal of `gt3x_comparison` is to compare reading in GT3X files from different devices and comparing the reading in from ActiLife and R/Python implementations.

## Reports

The [`old`](reports/old.html) shows the old NHANES GT3X format comparison. 

```{r}
x = list.files(pattern = ".Rmd", path = "reports", full.names = TRUE)
x = x[file.exists(sub("[.]Rmd$", ".html", x))]

```

## 
old_upper_limb

## Packages

The packages we'll use is:

```{r, echo = FALSE}
library(tibble)
df = tibble(
  url = c("https://github.com/THLfi/read.gt3x",
          "https://github.com/muschellij2/read.gt3x",
          "https://github.com/muschellij2/pygt3x",
          "https://github.com/muschellij2/SummarizedActigraphy",
          "https://github.com/paulhibbing/AGread"
  ),
  package = basename(url)
)
knitr::kable(df)
```


Examples of zeros in CSV from ActiLife: ID-PU13, serial-NEO1B41100262.  Still not giving exact answers: ID-PU2, serial-NEO1B41100255.

