---
title: "Analysis of Upper Limb Data Set"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the Packages

```{r load}
library(read.gt3x)
library(SummarizedActigraphy)
library(AGread)
library(dplyr)
library(pygt3x)
library(lubridate)
library(readxl)
library(dplyr)

if (packageVersion("read.gt3x") < "0.3.0") {
  stop(
    paste0(
      "Need to install new read.gt3x, ", 
      'remotes::install_github("muschellij2/read.gt3x")')
  )
}

if (packageVersion("pygt3x") < "0.0.8") {
  stop(
    paste0(
      "Need to install new pygt3x, ", 
      'remotes::install_github("muschellij2/pygt3x")')
  )
}

source(here::here("code/helper_functions.R"))
get_params = try(get("params"), silent = TRUE)
if (inherits(get_params, "try-error")) {
  params = list()
}
```

We are using the helper functions to read in the CSV files and download some data.

# Download the Data

## Downloading the Data


```{r get_fs_data}
## Personal dropbox sharing links to (1) GT3X file, (2) ActiLife raw data output CSV
gt3x.fpath <- "https://ndownloader.figshare.com/files/24642335"
csv.fpath <- "https://ndownloader.figshare.com/files/24642332"

file_directory <- tempdir()
gt3x_file <- file.path(file_directory, "TAS.gt3x.gz")
csv_file <- file.path(file_directory, "TAS.csv.gz")

## Download files to wd
if (!file.exists(gt3x_file)) download.file(gt3x.fpath, gt3x_file)
if (!file.exists(csv_file)) download.file(csv.fpath, csv_file)
```

# Read in the CSV

The ActiGraph CSV output is what we're considering as the "ground truth".  Note, the file format has a 10 line header and then the "data", and note that no time is given, it is assumed that the start date is the first record and then each record is 1 unit in the sampling frequency/sample rate (in this case 30Hz):

```{r}
readLines(csv_file, 15)
```

The `read_acc_csv` function simply reads in the header and the data (after skipping lines) and creates a `time` variable.  We want the column names to be `X`, `Y`, and `Z` so we strip off the `Accelerometer` label. 

```{r}
df = read_acc_csv(csv_file)
hdr = df$header
df = df$data
colnames(df) = sub("Accelerometer ", "", colnames(df))
head(hdr)
head(df)
tail(df)
nrow(df)
```

We see we have `r nrow(df)` records from the truth, which is not going to agree with the data below.  If the data does not agree, but all values are zero, I believe this is as good as we can assume to get.  This assumption is based on the fact we don't have the source code for the ActiLife software, and can only go off of https://github.com/actigraph/NHANES-GT3X-File-Format as a template.

# Read in the GT3X

## What is a GT3X
A `.gt3x` file is simply a zipped folder, but with specific files in there.  In the older NHANES format, we are interested in the `info.txt` and `activity.bin`.  In the newer format, we are interested in `info.txt` and `log.bin`.

```{r}
unzipped_gt3x = R.utils::gunzip(
  gt3x_file, remove = FALSE, temporary = TRUE, overwrite = TRUE)
tfile = tempfile()
output = unzip(unzipped_gt3x, exdir = tfile)
files = output[basename(output) %in% c("activity.bin", "log.bin", "info.txt")]
files
```

```{r unlink}
unlink(tfile, recursive = TRUE)
```


## read.gt3x

Here we are going to read in the GT3X file from `read.gt3x`:

```{r}
g = read.gt3x(gt3x_file, verbose = 2,
                   asDataFrame = TRUE)
xyz = c("X", "Y", "Z")
time_xyz = c("time", xyz)
act_df = read.gt3x(gt3x_file, verbose = 2,
                   asDataFrame = TRUE, imputeZeroes = TRUE)
head(act_df)
at = attributes(act_df)
at$light_data = NULL

class(act_df) = "data.frame"
act_df = act_df[, c("time", xyz)]
```


## AGread

Here we are going to read in the GT3X file from `AGread`, but it cannot at the time of running.  This is a check to see if new implementations/versions can read it in:
```{r}
ag = try({
  AGread::read_gt3x(gt3x_file, parser = "dev", flag_idle_sleep = TRUE)
}, silent = TRUE)
if (at$old_version) {
  stopifnot(inherits(ag, "try-error"))
  attributes(ag) = NULL; cat(ag)
} else {
  ag = ag$RAW
  colnames(ag) = trimws(sub("Accelerometer_", "", colnames(ag)))
  colnames(ag) = trimws(sub("Timestamp", "time", colnames(ag)))
  print(head(ag))
  print(table(ag$idle))
  ag = ag[, c("time", xyz, "idle")]
  round_away_zero = function (x, n = 0) {
    return(sign(x) * trunc(abs(x) * 10^n + 0.5)/10^n)
  }
  for (i in xyz) {
    ag[, i] = round_away_zero(ag[[i]], 3)
  }
}
```

## pygt3x

Here we are going to read in the GT3X file from `pygt3x` (https://github.com/muschellij2/pygt3x).  There are issues with the last values as we see here, which we do not recommend.  The `impute_zeros

```{r, error=TRUE}
py = pygt3x::py_read_gt3x(gt3x_file, verbose = TRUE)
max_py = max(abs(py$data)) 
py_success = max_py <= 8
stopifnot(py_success)
py = pygt3x::impute_zeros(py)
head(py)
tail(py)
```



## Comparison

Here we look at the records and see if they match up.  This comparison relies on a "good" reading of the data from `pygt3x` (which should now be fixed for old formats in versions `>= 0.0.8`, other testing needs to be done):

```{r}
if (py_success) {
  nrows = min(nrow(py), nrow(act_df))
  d = act_df[seq(nrows),] == py[seq(nrows),]
  stopifnot(all(d))
}
```

## "Fixing" zeros

In many cases there will be zeros for all values of `X`, `Y`, and `Z` axes.  There are some instances/measures you want to do the last observation carried forward, which is what idle sleep mode tends to do in ActiLife.

```{r}
act_df_fixed = SummarizedActigraphy::fix_zeros(act_df)
py_fixed = SummarizedActigraphy::fix_zeros(py)

# Not all Zeros are 
nrows = min(nrow(ag), nrow(act_df_fixed))
d = act_df_fixed[seq(nrows),time_xyz] == ag[seq(nrows),time_xyz]
bad = unname(rowSums(d) < ncol(d))
stopifnot(!all(d))

# Zeros are set to LOCF
ag_fixed = SummarizedActigraphy::fix_zeros(ag)
nrows = min(nrow(ag_fixed), nrow(act_df_fixed))
d = act_df_fixed[seq(nrows),time_xyz] == ag_fixed[seq(nrows),time_xyz]
bad = unname(rowSums(d) < ncol(d))
stopifnot(all(d))
```


## Full comparison

They agree, but only after last observation carried forward.

```{r}
df_equal = rowSums(df[, xyz] == act_df[, xyz]) == length(xyz)
stopifnot(all(act_df[!df_equal,xyz] == 0))

df2 = SummarizedActigraphy::fix_zeros(df)
stopifnot(all(df2[, xyz] == act_df_fixed[, xyz]))
stopifnot(all(df2[, xyz] == py_fixed[, xyz]))
```
