---
title: "Analysis of Upper Limb Data Set"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the Packages

```{r load}
library(read.gt3x)
library(SummarizedActigraphy)
library(AGread)
library(dplyr)
library(pygt3x)
library(lubridate)
library(readxl)
library(dplyr)

if (packageVersion("read.gt3x") < "0.3.0") {
  stop(
    paste0(
      "Need to install new read.gt3x, ", 
      'remotes::install_github("muschellij2/read.gt3x")')
  )
}

if (packageVersion("pygt3x") < "0.0.8") {
  stop(
    paste0(
      "Need to install new pygt3x, ", 
      'remotes::install_github("muschellij2/pygt3x")')
  )
}

source(here::here("code/helper_functions.R"))
get_params = try(get("params"), silent = TRUE)
if (inherits(get_params, "try-error")) {
  params = list()
}
```

We are using the helper functions to read in the CSV files and download some data.

# Download the Data

The data is from https://springernature.figshare.com/collections/Upper_limb_activity_of_twenty_myoelectric_prosthesis_users_and_twenty_healthy_anatomically_intact_adults_/4457855

We will be using a number of packages for this analysis.  We will use `dplyr`, `tidyr`, `readr` for data manipulation and input, `lubridate` to manipulate time (which is crucial with accelerometer data).  The `read.gt3x` package from https://github.com/THLfi/read.gt3x will allow us to read in `gt3x` files from ActiGraph.  The `AGread` package and `pygt3x` package can read in these files as well, but may have more limited functionality.  The `SummarizedActigraphy` (https://github.com/muschellij2/SummarizedActigraphy/) package is a wrapper for getting multiple accelerometry and wearable data types into a common data format.

# Data

The data is from @chadwell_kenney_granat_thies_galpin_head_2019.  The data consists of 40 subjects, 20 with prostheses, 20 without.  Each wore tri-axial ActiGraph watches for 7 days, one on each hand.  We will provide the demographics and clinical information related to this population below (in Section\@ref{metadata}).

The Figshare project for this data is located at https://springernature.figshare.com/collections/Upper_limb_activity_of_twenty_myoelectric_prosthesis_users_and_twenty_healthy_anatomically_intact_adults_/4457855. The gt3x data is located at https://springernature.figshare.com/articles/Unprocessed_raw_30Hz_acceleration_data_stored_as_gt3x/7946189.  This data consists of one directory of GT3X files.  In order to download them, you need to download the whole data set.  As the data are licensed under [CC0](https://creativecommons.org/publicdomain/zero/1.0/, we have copied the to another Figshare location (https://figshare.com/articles/dataset/Raw_GT3X_Data_Duplicate/11916087), where each `gt3x` file has been renamed without spaces or parentheses, each file can be downloaded separately, and each `gt3x` file has been compressed for faster transfer and smaller storage footprint.

Also, the raw CSVs from the ActiLife output is located at https://springernature.figshare.com/articles/dataset/Unprocessed_raw_30Hz_acceleration_data_stored_as_csv/7946186, which we have separated out into individual files (that line up with the GT3X files) at: https://figshare.com/articles/dataset/RAW_CSV_from_GT3X_Duplicate/12883463. 


## Downloading the Data

As the data is publicly available, you do not need to do any authentication to download the data.  You do, however, need to either 1) go to the original repository and download the full data set, 2) go to the Figshare page for the separated data and get the links for the files or download them via a browser, or 3) use the `rfigshare` package to obtain all the links to the data. 


```{r get_fs_data}
make_df = function(article_id) {
  x = rfigshare::fs_details(article_id = article_id, mine = FALSE,
                            session = NULL)
  files = x$files
  files = lapply(files, function(x) {
    as.data.frame(x[c("download_url", "name", "id", "size")],
                  stringsAsFactors = FALSE)
  })
  all_files = dplyr::bind_rows(files)
  all_files = all_files %>% 
    dplyr::filter(grepl("csv|gt3x", name)) %>% 
    tidyr::separate(name, into= c("id", "serial", "date"),
                    sep = "_", remove = FALSE) 
  all_files = all_files %>% 
    dplyr::mutate(
      group = gsub("\\d*", "", id),
      number = as.numeric(sub("AI|PU", "", id))
    ) %>% 
    dplyr::arrange(group, number) %>% 
    tibble::as_tibble()
}
gt3x = make_df("11916087") %>% 
  dplyr::select(name, download_url, id, serial)
csv = make_df("12883463") %>% 
  dplyr::select(name, download_url, id, serial)

data_df = dplyr::full_join(
  gt3x, csv, 
  by = c("id", "serial"),
  suffix = c("_gt3x", "_csv"))
stopifnot(nrow(data_df) == nrow(gt3x))
stopifnot(nrow(data_df) == nrow(csv))
nrow(csv)
```


```{r}
# irow = sample(nrow(data_df), 1)
irow = params$irow
if (is.null(irow)) {
  irow = 5
}
print(irow)
idf = data_df[irow,]
print(idf)
download = function(url, name) {
  destfile = file.path(tempdir(), name)
  if (!file.exists(destfile)) {
    download.file(url, destfile)
  }
  destfile
}
gt3x_file = download(idf$download_url_gt3x, idf$name_gt3x)
csv_file = download(idf$download_url_csv, idf$name_csv)
```

# Read in the CSV

The ActiGraph CSV output is what we're considering as the "ground truth".  Note, the file format has a 10 line header and then the "data", and note that no time is given, it is assumed that the start date is the first record and then each record is 1 unit in the sampling frequency/sample rate (in this case 30Hz):

```{r}
readLines(csv_file, 15)
```

The `read_acc_csv` function simply reads in the header and the data (after skipping lines) and creates a `time` variable.  We want the column names to be `X`, `Y`, and `Z` so we strip off the `Accelerometer` label. 

```{r}
df = read_acc_csv(csv_file)
hdr = df$header
df = df$data
colnames(df) = sub("Accelerometer ", "", colnames(df))
head(hdr)
head(df)
tail(df)
nrow(df)
```

We see we have `r nrow(df)` records from the truth, which is not going to agree with the data below.  If the data does not agree, but all values are zero, I believe this is as good as we can assume to get.  This assumption is based on the fact we don't have the source code for the ActiLife software, and can only go off of https://github.com/actigraph/NHANES-GT3X-File-Format as a template.

# Read in the GT3X

## What is a GT3X
A `.gt3x` file is simply a zipped folder, but with specific files in there.  In the older NHANES format, we are interested in the `info.txt` and `activity.bin`.  In the newer format, we are interested in `info.txt` and `log.bin`.

```{r}
unzipped_gt3x = R.utils::gunzip(
  gt3x_file, remove = FALSE, temporary = TRUE, overwrite = TRUE)
tfile = tempfile()
output = unzip(unzipped_gt3x, exdir = tfile)
files = output[basename(output) %in% c("activity.bin", "log.bin", "info.txt")]
files
```

```{r unlink}
unlink(tfile, recursive = TRUE)
```


## read.gt3x

Here we are going to read in the GT3X file from `read.gt3x`:

```{r}
act_df = read.gt3x(gt3x_file, verbose = TRUE,
                   asDataFrame = TRUE, imputeZeroes = TRUE)
head(act_df)
at = attributes(act_df)
at$light_data = NULL

class(act_df) = "data.frame"
act_df = act_df[, c("time", "X", "Y", "Z")]
```


### Errors in Time

Here we see the last time of the data set is **after** the download date of the data - indicating some errors in the data.  This is referenced in https://github.com/THLfi/read.gt3x/issues/20 and https://github.com/THLfi/read.gt3x/issues/3, which are issues that are in the works of hopefully being fixed.

```{r}
print(at$header$`Sample Rate`)
at$header$`Download Date`
tail(act_df)
last_time = act_df$time[ nrow(act_df)]

if (at$old_version) {
  bad_time = last_time > at$header$`Download Date`
  bad_time = bad_time | (last_time > at$header$`Stop Date`)
  print(bad_time)
  stopifnot(bad_time)
}
```

In future releases of `read.gt3x`, this will stop this report to indicates changes need to be made:

```{r}
if (at$old_version) {
  stopifnot(bad_time)
}
```

Right now, we will make the time ourselves

```{r}
act_df$time = at$start_time +
  lubridate::as.period(
    (0:(nrow(act_df)-1))/at$sample_rate,
    unit = "secs")

last_time2 = act_df$time[ nrow(act_df)]
last_time2 == at$header$`Download Date`
abs((last_time2 - at$header$`Stop Date`)) < 0.5
```

## AGread

Here we are going to read in the GT3X file from `AGread`, but it cannot at the time of running.  This is a check to see if new implementations/versions can read it in:
```{r}
ag = try({
  AGread::read_gt3x(gt3x_file, parser = "dev")
}, silent = TRUE)
if (at$old_version) {
  stopifnot(inherits(ag, "try-error"))
  attributes(ag) = NULL; cat(ag)
} else {
  ag = ag$RAW
  colnames(ag) = trimws(sub("Accelerometer_", "", colnames(ag)))
  colnames(ag) = trimws(sub("Timestamp", "time", colnames(ag)))
  ag = ag[, c("time", "X", "Y", "Z")]
}
```

## pygt3x

Here we are going to read in the GT3X file from `pygt3x` (https://github.com/muschellij2/pygt3x).  There are issues with the last values as we see here, which we do not recommend.  The `impute_zeros

```{r, error=TRUE}
py = pygt3x::py_read_gt3x(gt3x_file)
max_py = max(abs(py$data)) 
py_success = max_py < 6.1
stopifnot(py_success)
py = pygt3x::impute_zeros(py)
head(py)
tail(py)
```



## Comparison

Here we look at the records and see if they match up.  This comparison relies on a "good" reading of the data from `pygt3x` (which should now be fixed for old formats in versions `>= 0.0.8`, other testing needs to be done):

```{r}
if (py_success) {
  nrows = min(nrow(py), nrow(act_df))
  d = act_df[seq(nrows),] == py[seq(nrows),]
  stopifnot(all(d))
}
```

## "Fixing" zeros

In many cases there will be zeros for all values of `X`, `Y`, and `Z` axes.  There are some instances/measures you want to do the last observation carried forward, which is what idle sleep mode tends to do in ActiLife.

```{r}
act_df_fixed = SummarizedActigraphy::fix_zeros(act_df)
py_fixed = SummarizedActigraphy::fix_zeros(py)
```


## Full comparison

They agree!

```{r}
stopifnot(all(df[, c("X", "Y", "Z")] == act_df_fixed[, c("X", "Y", "Z")]))
stopifnot(all(df[, c("X", "Y", "Z")] == py_fixed[, c("X", "Y", "Z")]))

stopifnot(all(df[, c("X", "Y", "Z")] == act_df[, c("X", "Y", "Z")]))
stopifnot(all(df[, c("X", "Y", "Z")] == py[, c("X", "Y", "Z")]))
```
